---
title: "Volleyball Waffle Chart"
format: html
---

```{r}
library(tidyverse)
library(waffle)

matches <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensVolleyballData/main/data/ncaa_womens_volleyball_matchstats_2025.csv")

md <- matches |> 
  filter(team == "Maryland") |> 
  summarise(
    kills = sum(kills), 
    blocks = sum(block_solos + block_assists), 
    aces = sum(aces), 
    opponent_errors = sum(errors)
  )

mi <- matches |>  
filter(team == "Michigan") |> 
  summarise(
    kills = sum(kills), 
    blocks = sum(block_solos + block_assists), 
    aces = sum(aces), 
    opponent_errors = sum(errors)
  )
```
```{r}
md <- c("kills"=525, "blocks"=216, "aces"=72, "opponent_errors"=247)
mi <- c("kills"=760, "blocks"=221, "aces"=82, "opponent_errors"=243)
```



```{r}
  waffle(md,
         rows = 10,
         title = "Maryland Volleyball: Points by Source",
         xlab = "1 square = 25 points",
         colors = c("red", "black", "gold", "gray"))
```
```{r}
  waffle(mi,
         rows = 10,
         title = "Michigan Volleyball",
         xlab = "1 square = 25 points",
         colors = c("red", "black", "gold", "gray"))
```

Maryland's offense this season has been heavily driven by kills, which accounted for 525 total points.They also picked up 247 points from opponent errors, showing that forcing mistakes plays a key role in their scoring. Blocks and aces made smaller contributions reflecting a balanced appraoch between defense and pressure. On the other hand, Michigan also relief on kills with 760 showing that they have a more dominant offensive position. Their numbers in blocks and aces were similar to Maryland's but they benefited less from overall opponent errors. 
