---
title: "Coaching Networks and Career Pathways Analysis"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(igraph)
library(ggraph)
library(tidygraph)
library(lubridate)
```

# Overview

This analysis explores coaching networks and career pathways in NCAA women's basketball, focusing on which schools produce future head coaches and identifying coaching "trees."

Your Task: Work through each section, making decisions and identifying potential story angles from the data.

---

# Data Loading

Load the CSV file and check out the data to see what you're working with.

```{r load-data}
coaching <- read_csv("https://raw.githubusercontent.com/dwillis/wbb/refs/heads/main/ncaa/coaches/coaching_histories_merged.csv")

cat("\nDataset summary:\n")
cat("Total records:", nrow(coaching), "\n")
cat("Unique coaches:", n_distinct(coaching$coach), "\n")
cat("Unique schools:", n_distinct(coaching$college), "\n")
cat("Division I records:", sum(coaching$division == "I", na.rm = TRUE), "\n")
```

---

# Analysis 1: Head Coach Production Pipeline

## DECISION POINT 1: Defining "Future Head Coaches"

**Question for you**: Should we focus only on coaches who became D1 head coaches, or also include those who became head coaches at D2, D3, or other levels? How does this choice affect the story? Does it matter if someone became a D3 head coach vs. stayed in D1 as an assistant?

I think it is best to focys on coaches who ended up as D1 head coaches so the story stays clear and shows the path that lead to the highest level of women's college basketball. Including D2 or D3 head coaches could be interesting, but thats combines different career paths and can make comparisons harder. Being a D3 head coach still shows leadership, and staying in D1 as an assistant usually comes with more prestige. Focusing on D1 head coaches would let us see real patterns and pipelines and you can always end up mentioning lower division paths shortly to show alternative routes. 


**Your choice** (modify the code below):

```{r head-coach-pipeline}
# STUDENT DECISION: Modify this filter as needed
# Option A: Only D1 head coaches (current setting)
# Option B: All NCAA head coaches (remove division filter)
# Option C: D1 + D2 only (division %in% c("I", "II"))

d1_head_coaches <- coaching |>
  filter(division == "I", position_title_standardized == "Head Coach") |> 
  distinct(coach) |>
  pull(coach)

cat("Number of head coaches identified:", length(d1_head_coaches), "\n")
```


## Finding Previous Positions

```{r previous-positions}
# Find their previous positions (before becoming head coach)
future_hc_positions <- coaching |>
  filter(coach %in% d1_head_coaches, position_title_standardized != 'Head Coach') |>
  arrange(coach, start_year) |>
  group_by(coach) |>
  mutate(
    is_head_coach = position_title_standardized == "Head Coach",
    first_hc_year = min(start_year[is_head_coach], na.rm = TRUE),
    before_hc = start_year < first_hc_year | is.na(first_hc_year)
  ) |>
  ungroup() 

cat("Total positions held by future head coaches:", nrow(future_hc_positions), "\n")
```

## DECISION POINT 2: Which Division Matters for Training?

**Question for you**: When looking at which schools produce future head coaches, should we only count their D1 experience, or also their D2/D3/other experience?

I think it depends on the scenario. If we were trying to highlight timelines and pathways to the higest level, it would make sense to only count their Dq experience so that were looking at schools that directly feed into Division I head coaching jobs. Using D2, D3, or other experience can be interesting for context, since it shows how coaches build leadership experiences in those lower divisons, but it can also be unneeded information if were only trying to focus on D1 pathways
```{r production-schools}
# STUDENT DECISION: Modify the division filter
# Current: Only counting D1 positions
# Alternative: Remove division filter to count all positions

hc_production_schools <- future_hc_positions |>
  filter(before_hc, division == "I") |>  
  group_by(college_clean, position_title_standardized) |>
  summarize(
    future_hcs = n_distinct(coach),
    .groups = "drop"
  ) |>
  arrange(desc(future_hcs))

cat("\nTop 20 D1 schools producing future head coaches (all positions):\n")
hc_production_schools |>
  group_by(college_clean) |>
  summarize(total_future_hcs = sum(future_hcs), .groups = "drop") |>
  arrange(desc(total_future_hcs)) |>
  head(20) |>
  print(n = 20)
```

**Think about**: Does a stint at a D3 school early in someone's career "count" as part of their development? 

I would say yes, probably, it gives coaches hands on experience. 

## Position Type Breakdown

```{r position-breakdown}
cat("\n\nTop schools by position type:\n")
hc_production_schools |>
  filter(position_title_standardized %in% c("Assistant Coach", "Associate Head Coach")) |>
  group_by(position_title_standardized) |>
  slice_max(future_hcs, n = 10) |>
  arrange(position_title_standardized, desc(future_hcs))
```

### 💡 STORY IDEAS - Section 1

**Based on what you see above, brainstorm 2-3 story angles:**

1. "Pipeline Powerhouses" this would include focusing on schools like Mississippi State and Wake Forest, which produce the most future head coaches. It would be interesting to explore what makes these orgrams consistent deployers and determine if it is majority networking opportunities
2. "Mid-Major Movers" this could highlight mid programs and discuss how these schools provide unique oppurtunities for assistants to fevelop leadearship skills and move up the ranks
3. "Regional Coaching Trends" this would look at whether certain regions or areas dominate the production of future head coaches. Are schools in the Southeast or Northeast disproportionately represented? And we would tie it to recruiting patterns, conference strength, and historical coaching networks  

**Questions to consider:**
- Which school dominates the list? Is this surprising?
- Are there mid-major schools punching above their weight?
- Are there regional patterns?

---

# Analysis 2: Career Ladder Analysis

## DECISION POINT 3: Minimum Position Threshold

**Question for you**: Should we require a minimum number of positions before someone counts as having a "career ladder"? (e.g., exclude people who went straight from 1 assistant job to head coach?)

Yes. I think it would make sense to set a minimum threshold for defining a "career ladder," especially if we want average pathways and developmet patterns. Prior 2-3 positions should be good

```{r career-ladder}
# STUDENT DECISION: Set minimum number of positions
MIN_POSITIONS <- 1  # CHANGE THIS NUMBER (try 2, 3, or keep at 1)

# Sequence of positions before becoming head coach
career_sequences <- future_hc_positions |>
  filter(before_hc, division == "I") |>
  arrange(coach, start_year) |>
  group_by(coach) |>
  summarize(
    positions_held = list(position_title_standardized),
    n_positions = n(),
    final_position = last(position_title_standardized),
    .groups = "drop"
  ) |>
  filter(n_positions >= MIN_POSITIONS)  # APPLIES YOUR THRESHOLD

cat("Coaches meeting threshold:", nrow(career_sequences), "\n")

cat("\nMost common final position before becoming D1 head coach:\n")
career_sequences |>
  count(final_position, sort = TRUE) |>
  mutate(pct = n / sum(n) * 100) |>
  head(10)

# Average number of positions held before head coach role
cat("\n\nAverage D1 positions held before becoming head coach:",
    round(mean(career_sequences$n_positions, na.rm = TRUE), 1))
```

## Common Career Pathways

```{r pathways}
# STUDENT DECISION: How many positions should we examine?
PATHWAY_LENGTH <- 3  

cat("\n\nCommon", PATHWAY_LENGTH, "-position pathways to head coach:\n")
career_sequences |>
  filter(n_positions == PATHWAY_LENGTH) |>
  mutate(pathway = map_chr(positions_held, ~paste(.x, collapse = " → "))) |>
  count(pathway, sort = TRUE) |>
  head(15) |>
  print(n = 15)
```

### 💡 STORY IDEAS - Section 2

**Based on the career ladder data, brainstorm 2-3 story angles:**

1. "The Classic Climb" this would highlught the most common pathways and explore why spending multiple years as an assistant is the standard route, how experience accumulates, and what skills are gained at each step before reaching head coach status
2. "The Fast Trackers" would focus on head coaches who have skipped steps, for example, moving from asistant to interm head coach or associate head coach to head coach. We could look at what factors allow some coaches to accelerate their careers
3. "Staff as a Launchpad" to examine how starting as staff positions contributes to head coaching roles. We can look at cases where coaches start in non assistant roles and work their way up

**Questions to consider:**
- Are there "fast track" coaches who skipped steps?
- Does the Associate Head Coach position matter, or do people jump from Assistant to Head?
- How important is the "Staff" level as a starting point?

---

# Analysis 3: Coaching Trees and Mentorship Networks

## DECISION POINT 4: Time Overlap Requirements

**Question for you**: How should we define "worked under"? Should we require they overlapped for at least one full year, or is any overlap sufficient?

We should probably require at leasy one full season of overlap. This ensures the assitant or staff member actually had meaninful exposure to the head coaches leadership and decision making. Using an overlap could overstate connections and make it seem like coaches "worked under" someone when the experience was minimal 
```{r coaching-trees}
# STUDENT DECISION: Set minimum overlap threshold
MIN_OVERLAP_YEARS <- 1  

# Create edges: assistant coaches who worked under head coaches
coaching_relationships <- coaching |>
  filter(division == "I") |>
  select(coach, college_clean, position_title_standardized, start_year, end_year) |>
  # For each school/timeframe, find head coaches
  left_join(
    coaching |>
      filter(division == "I", position_title_standardized == "Head Coach") |>
      select(hc_name = coach, college_clean, 
             hc_start = start_year, hc_end = end_year),
    by = "college_clean",
    relationship = "many-to-many"
  ) |>
  # Keep only overlapping tenures (assistant worked under this head coach)
  filter(
    !is.na(hc_start),
    coach != hc_name,
    position_title_standardized %in% c("Assistant Coach", "Associate Head Coach"),
    # Check for time overlap
    start_year <= coalesce(hc_end, 2025),
    coalesce(end_year, 2025) >= hc_start
  ) |>
  # STUDENT DECISION: Add minimum overlap requirement
  mutate(
    overlap_years = pmin(coalesce(end_year, 2025), coalesce(hc_end, 2025)) - 
                    pmax(start_year, hc_start)
  ) |>
  filter(overlap_years >= MIN_OVERLAP_YEARS) |>  # APPLIES YOUR THRESHOLD
  distinct(coach, hc_name, college_clean)

cat("Coaching relationships identified:", nrow(coaching_relationships), "\n")
```

## Top Mentors

```{r mentors}
# Most prolific mentors (head coaches who produced assistants who became HCs)
prolific_mentors <- coaching_relationships |>
  filter(coach %in% d1_head_coaches) |>
  count(hc_name, sort = TRUE) |>
  rename(mentor = hc_name, proteges_who_became_hc = n)

cat("\nTop 20 mentors (head coaches whose assistants became D1 head coaches):\n")
print(prolific_mentors |> head(20), n = 20)
```

### 💡 STORY IDEAS - Section 3

**Based on the coaching trees data, brainstorm 2-3 story angles:**

1. "The Mentorship Giants" this would focus on strong mentors like Brenda Frese and Kim Barnes Arico, who have sent multiple asistants on to D1 head coaching roles. We could explore what makes their programs effective at developing future leaders
2. "Unexpected Influence" could highlight names on the list that are surprising like Wes Moore or Mike Bradbury, and look at how they've shaped future head coaches.
3. "Mapping a Coaching Tree" In this we would pick a specific mentor like Dawn Staley or Brenda Freese, and track their progress accross multiple schools. This story could visualize a coaching family tree showing how one's leader's influeunce spreads through women's college basketball 

**Questions to consider:**
- Who are the most prolific "mentor" coaches?
- Are there surprising names?
- Do certain coaches have "coaching trees" at multiple schools?
- Could you profile a specific coaching tree? Which one?

## DECISION POINT 5: Network Visualization Scope

**Question for you**: How many top mentors should we include in the visualization? Too many = cluttered, too few = missing important connections.

10-15 is good.

```{r network-visualization, fig.width=12, fig.height=10}
# STUDENT DECISION: Number of top mentors to visualize
N_MENTORS_TO_SHOW <- 15  # CHANGE THIS (try 10, 15, 20, or 25)

# Create network of top mentors (for visualization clarity)
top_mentors <- prolific_mentors |>
  head(N_MENTORS_TO_SHOW) |>
  pull(mentor)

network_edges <- coaching_relationships |>
  filter(
    hc_name %in% top_mentors,
    coach %in% d1_head_coaches
  ) |>
  select(from = hc_name, to = coach, school = college_clean)

# Build network graph
coaching_network <- as_tbl_graph(network_edges, directed = TRUE)

# Plot
ggraph(coaching_network, layout = 'fr') +
  geom_edge_link(aes(alpha = 0.3), arrow = arrow(length = unit(2, 'mm'))) +
  geom_node_point(size = 5, color = "steelblue") +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  theme_void() +
  labs(
    title = "Coaching Trees: Top Mentors and Their Protégés",
    subtitle = paste("Top", N_MENTORS_TO_SHOW, "D1 head coaches who mentored future D1 head coaches")
  )
```

---

# Analysis 4: School-to-School Coaching Pipelines

```{r school-pipelines}
# Track coach movement between schools
coach_movements <- coaching |>
  filter(division == "I") |>
  arrange(coach, start_year) |>
  group_by(coach) |>
  mutate(
    next_school = lead(college_clean),
    next_position = lead(position_title_standardized),
    next_start = lead(start_year)
  ) |>
  filter(!is.na(next_school), college_clean != next_school) |>
  ungroup()

# Most common school-to-school transitions
cat("\nTop 30 school-to-school coaching pipelines:\n")
coach_movements |>
  count(college_clean, next_school, sort = TRUE) |>
  rename(from_school = college_clean, to_school = next_school, coaches = n) |>
  head(30) |>
  print(n = 30)
```

## Schools that Export to Head Coaching

```{r exports}
# Schools that "export" coaches to D1 head coaching jobs
coach_exports <- coach_movements |>
  filter(next_position == "Head Coach") |>
  count(college_clean, sort = TRUE) |>
  rename(school = college_clean, coaches_to_hc = n)

cat("\n\nTop 20 schools that export coaches to D1 head coaching positions:\n")
print(coach_exports |> head(20), n = 20)
```

### 💡 STORY IDEAS - Section 4

**Based on school-to-school movement patterns, brainstorm 2-3 story angles:**

1. "Regional Powerhouse and Pipelines" we could explore strong conference based pipelines like Memphis -> Cincinnati or Mississippi State -> Texas to show how certain regons dominate the movement of coaching talent.
2. "From Top Programs to Mid-Majors" to focus on how assistants at high profile schools often get their first head coaching jobs at smaller programs like Wake Forest -> Davidson
3. "Unexpected Connections" to highlught unusual or cross country pipelines, like Missouri St. to Virginia 

**Questions to consider:**
- Are there strong regional pipelines (e.g., schools from the same conference feeding each other)?
- Do assistant coaches at top programs tend to move to head coaching at mid-majors first?
- Are there unusual cross-country connections?
- Which schools are "feeders" vs. "destinations"?

---

# Analysis 5: Conference-Level Analysis

## DECISION POINT 6: Conference Focus

**Question for you**: Should we analyze all conferences equally, or focus on major/Power 5 conferences? Should we combine some smaller conferences?

Focusing on all conferences would allow us to have a full clear picture of the high level coaching pipelines. 

```{r conference-analysis}
# STUDENT DECISION: Define which conferences to highlight
# Option A: Analyze all conferences (current)
# Option B: Create a "Power 5" vs "Mid-Major" vs "Small" grouping
# Option C: Focus only on specific conferences

# Conferences producing future head coaches
conference_production <- future_hc_positions |>
  filter(before_hc, division == "I") |>
  count(conference, position_title_standardized) |>
  arrange(desc(n))

cat("\nTop conferences producing future D1 head coaches:\n")
conference_production |>
  group_by(conference) |>
  summarize(total = sum(n), .groups = "drop") |>
  arrange(desc(total)) |>
  head(15) |>
  print(n = 15)
```

## Conference Loyalty

```{r loyalty}
# Conference loyalty: coaches who stay within conference
conference_loyalty <- coach_movements |>
  left_join(
    coaching |> 
      filter(division == "I") |>
      select(coach, college_clean, conference),
    by = c("next_school" = "college_clean", "coach" = "coach"),
    relationship = "many-to-many"
  ) |>
  filter(!is.na(conference.x), !is.na(conference.y)) |>
  mutate(same_conference = conference.x == conference.y) |>
  group_by(conference.x) |>
  summarize(
    total_moves = n(),
    same_conference_moves = sum(same_conference),
    loyalty_rate = same_conference_moves / total_moves * 100,
    .groups = "drop"
  ) |>
  arrange(desc(loyalty_rate))

cat("\n\nConference retention rates (coaches staying within conference):\n")
print(conference_loyalty |> head(15), n = 15)
```

### 💡 STORY IDEAS - Section 5

**Based on conference patterns, brainstorm 2-3 story angles:**

1. "Staying Loyal vs Moving On" could examine loyalty rates. Some smaller conferences like the SWAC and Big Sky, show higher retention, while Power 5 conferences have lower loyalty. This could help create a story on why coaches move out of top conferences
2. "Importers and Exporters" would look at which conferences produce coaches who move somewhere else versus hiring coaches from other conferences 
3. "Coach Factories of Women's Basketball" to highlught conferences that produce the most future D1 head coaches, like the ACC, SEC, and Atlantic 10. We would explore what makes conferences effective at developing coaching talent 

**Questions to consider:**
- Do certain conferences "develop" more coaches than others?
- Is there high loyalty within conferences, or do coaches move around?
- Are there conferences that import talent vs. export it?

---

# Analysis 6: Time to Head Coach Role

## DECISION POINT 7: Career Start Definition

**Question for you**: When calculating "time to head coach," should we count ALL positions (including high school, community college) or only college-level positions?

It's best to count only college level positions for "time to head coach" because including high school or community college roles could skew comparisons

```{r time-to-hc}
# STUDENT DECISION: Filter on category
# Current: All positions included
# Alternative: Add filter like "category == 'College'" to exclude high school, etc.

time_to_hc <- coaching |>
  filter(coach %in% d1_head_coaches) |>
  # filter(category == "College") |>
  arrange(coach, start_year) |>
  group_by(coach) |>
  summarize(
    first_position_year = min(start_year, na.rm = TRUE),
    first_hc_year = min(start_year[position_title_standardized == "Head Coach" & 
                                     division == "I"], na.rm = TRUE),
    years_to_hc = first_hc_year - first_position_year,
    .groups = "drop"
  ) |>
  filter(is.finite(years_to_hc), years_to_hc >= 0)

cat("\nTime to D1 head coach position:\n")
cat("Mean:", round(mean(time_to_hc$years_to_hc, na.rm = TRUE), 1), "years\n")
cat("Median:", median(time_to_hc$years_to_hc, na.rm = TRUE), "years\n")
cat("Min:", min(time_to_hc$years_to_hc, na.rm = TRUE), "years\n")
cat("Max:", max(time_to_hc$years_to_hc, na.rm = TRUE), "years\n")

# Distribution
ggplot(time_to_hc, aes(x = years_to_hc)) +
  geom_histogram(binwidth = 2, fill = "steelblue", alpha = 0.7) +
  geom_vline(xintercept = median(time_to_hc$years_to_hc, na.rm = TRUE), 
             linetype = "dashed", color = "red", size = 1) +
  labs(
    title = "Years to First D1 Head Coaching Position",
    subtitle = "From first coaching role to D1 head coach",
    x = "Years",
    y = "Count of coaches"
  ) +
  theme_minimal()
```

### 💡 STORY IDEAS - Section 6

**Based on career timeline data, brainstorm 2-3 story angles:**

1. "The Typical CLimb:" Explore that the average coach takes about 12 years to reach a D1 head coaching position, highlighting the loing term development and experience needed
2. "Fast Movers vs Late Bloomers": Focys on outliers, like coaches who reached D1 head coach in 0-5 years versus those who took 30+ years
3. "The Long Game of Coaching:" Talk about how even with rapid rises, most coaches spend a decade gaining experience as assitants or at lower divisons to emphasize the patience and persistence behind building a career in women's college basketball 

**Questions to consider:**
- What's the typical timeline to becoming a D1 head coach?
- Who are the outliers (very fast or very slow)?

---

# Analysis 7: Elite Program "Finishing Schools"

## DECISION POINT 8: Defining "Elite Programs"

**Question for you**: Which schools should count as "elite" or "blue blood" programs? 
This is subjective - modify the list based on your knowledge or research.

Elite or blue blood programs are schools with a history of success, national championships, or consistent NCAA tournament shows

```{r finishing-schools}
# STUDENT DECISION: Modify this list of top programs
# Consider: recent success, historical success, resources, national profile

top_programs <- c("UConn", "Tennessee", "Stanford", "Notre Dame", "Baylor",
                  "South Carolina",
                  "Louisville", "Louisiana Tech")
# ADD OR REMOVE SCHOOLS AS YOU SEE FIT

cat("Analyzing", length(top_programs), "programs\n")

# Assistant coaches from top programs who became head coaches
finishing_school_analysis <- future_hc_positions |>
  filter(
    before_hc,
    division == "I",
    college_clean %in% top_programs,
    position_title_standardized %in% c("Assistant Coach", "Associate Head Coach")
  ) |>
  group_by(college_clean) |>
  summarize(
    future_head_coaches = n_distinct(coach),
    .groups = "drop"
  ) |>
  arrange(desc(future_head_coaches))

cat("\n'Finishing schools' - top programs producing future head coaches:\n")
print(finishing_school_analysis, n = Inf)

# Average years spent at top program before becoming HC
years_at_top_programs <- future_hc_positions |>
  filter(
    before_hc,
    division == "I",
    college_clean %in% top_programs,
    position_title_standardized %in% c("Assistant Coach", "Associate Head Coach")
  ) |>
  mutate(years_at_school = coalesce(end_year, 2025) - start_year) |>
  group_by(college_clean) |>
  summarize(
    avg_years = mean(years_at_school, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(avg_years))

cat("\n\nAverage years spent at top programs (for future HCs):\n")
print(years_at_top_programs, n = Inf)
```

### 💡 STORY IDEAS - Section 7

**Based on elite program patterns, brainstorm 2-3 story angles:**

1. "Short Term vs Long Tenures:" Examine the length of time spent at elite programs and how it affects career outcomes
2. "Hidden Gems": Focus on programs that pull their weught like Louisiana Tech or Stanford, and look at whether spending more years at an elite program correlates with future coaching success
3. "The Finishing Schools:" Highlight top programs like Louisville, Tennessee, and South Carolina that produce future head coaches and explore what makes them so effective 

**Questions to consider:**
- Do the biggest brand names produce the most head coaches?
- Are there "hidden gems" - programs that punch above their weight?
- Does time spent at elite programs matter (short stint vs. long apprenticeship)?
- Are there programs that hoard talent vs. those that develop and release it?

---

# Analysis 8: "Stepping Stone" Schools

```{r stepping-stones}
# Head coaches who left for another D1 head coaching job
stepping_stone_moves <- coaching |>
  filter(division == "I", position_title_standardized == "Head Coach") |>
  arrange(coach, start_year) |>
  group_by(coach) |>
  mutate(
    next_school = lead(college_clean),
    next_position = lead(position_title_standardized),
    moved_to_hc = next_position == "Head Coach" & !is.na(next_school)
  ) |>
  filter(moved_to_hc) |>
  ungroup()

cat("\nSchools most commonly serving as stepping stones (HC → HC moves):\n")
stepping_stone_moves |>
  count(college_clean, sort = TRUE) |>
  rename(school = college_clean, hc_departures = n) |>
  head(20) |>
  print(n = 20)

# Most common destination schools
cat("\n\nMost common destination schools for head coaches:\n")
stepping_stone_moves |>
  count(next_school, sort = TRUE) |>
  rename(destination = next_school, hc_arrivals = n) |>
  head(20) |>
  print(n = 20)
```

### 💡 STORY IDEAS - Section 8

**Based on stepping stone patterns, brainstorm 2-3 story angles:**

1. "Destinations vs Departures": Compare schools that consistently retain coaches versus those that lose coaches quickly and explore why programs like GW or Texas Texh attract head coaches
2. "The Launch Pads": Highlight schools that frequently serve as a first head coaching job
3. "Hidden Pipelines": Focus on lesser known programs that feed coaches into other schools showing that coaching development isnt limited to big name programs 

**Questions to consider:**
- Which schools can't keep their successful head coaches?
- What makes a school a "destination" vs. a "stepping stone"?
- Are there schools that serve as consistent "first head coaching jobs"?
- Can you identify a "ladder" of schools (small to medium to major)?

---

# Analysis 9: Position Mobility Analysis

```{r position-mobility}
# Years spent in each position before moving up
position_duration <- coaching |>
  filter(division == "I") |>
  mutate(years_in_position = coalesce(end_year, 2025) - start_year) |>
  group_by(position_title_standardized) |>
  summarize(
    n_positions = n(),
    avg_duration = mean(years_in_position, na.rm = TRUE),
    median_duration = median(years_in_position, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(n_positions))

cat("\nAverage years in each position type:\n")
print(position_duration, n = Inf)

# Promotion rates by position
promotions <- coaching |>
  filter(division == "I") |>
  arrange(coach, start_year) |>
  group_by(coach) |>
  mutate(
    next_position = lead(position_title_standardized),
    promoted = case_when(
      position_title_standardized == "Staff" & 
        next_position %in% c("Assistant Coach", "Associate Head Coach", "Head Coach") ~ TRUE,
      position_title_standardized == "Assistant Coach" & 
        next_position %in% c("Associate Head Coach", "Head Coach") ~ TRUE,
      position_title_standardized == "Associate Head Coach" & 
        next_position == "Head Coach" ~ TRUE,
      TRUE ~ FALSE
    )
  ) |>
  filter(!is.na(next_position)) |>
  group_by(position_title_standardized) |>
  summarize(
    total_moves = n(),
    promotions = sum(promoted),
    promotion_rate = promotions / total_moves * 100,
    .groups = "drop"
  ) |>
  arrange(desc(promotion_rate))

cat("\n\nPromotion rates by position:\n")
print(promotions, n = Inf)
```

### 💡 STORY IDEAS - Section 9

**Based on mobility and promotion patterns, brainstorm 2-3 story angles:**

1. "Climbing the Ladder": Map typical progress paths from staff -> Assistant -> Associate -> Head Coach, showing how mobility differs by position and how career planning influences the speed of advancement 
2. "Springboard vs Dead Ends": Highlught positions like Staff and Associate Head Coach as stepping stones vs roles like Head Coach, Interim Head Coach, and Recruiting Coordinator which rarely leads to further promotion 
3. "Timing Matters": Examine how long coaches typically stay in each tole and duscuss whetther there is an optimal time that needs to be spend in that position before moving up

**Questions to consider:**
- Which positions are "dead ends" vs. "springboards"?
- Do people spend more time in certain positions?
- Is there a "right" amount of time to spend before moving up?

---

# FINAL REFLECTION: Your Best Story Ideas

Review all your story ideas from each section. Which 3 would you pursue as full stories?

**Story #1:** "Pipeline Powerhouses: Schools That Produce D1 Head Cocaches" 

**Why this story:**
This story would focus on programs like Mississippi State, Wake Forest, and other schools that consistently produce future head coaches. It explores what makes these programs effective at developing coaching talent, networking, mentorship, and leadership oppuurtunities. It would appeal to readers interested in the developmental side of college basketball and gives concrete examples of programs shaping the sport beyond wins and losses
---

**Story #2:**
"The Classic Climb vs Fast Trackers: Career Ladders in Women's College Basketball" 
**Why this story:**
This story compares the typical path to becoming a D1 head coach versus outliers who accelerated quickly or took unusually long. It highlughts the steps, positions, and average years spend at each role, giving insight into career planning, mentorship, and oppurtunity structures in NCAA basketball 
---

**Story #3:**
"Mentorship Giants and Coaching Trees: Mapping Influence in Women's Basketball" 
**Why this story:**
This story profiles the strongest mentors like Brenda Frese showing how their proteges move on to become D1 head coaches. Using network visualizations, we can map coaching trees to show influence and conenctivity across programs. This story could be visually engaging and data driven while shedding a light on the social and professional network that shape women's college basketball careers
---

# Additional Analysis Ideas to Explore

Now that you've worked through these analyses, what other questions could this data answer?

**Your ideas:**

1. Look at whether coaches who start as Staff vs Assistant Coaches reach coaching positions faster or slower 
2. Analyze gender or diversity patterns among mentors and proteges to see if certain programs or coaching trees are more inclusive
3. Examine outliers in career timelines, like coaches who reached D1 head coach quickly or took unusually long 