## NCAA Football Attendance 

Let's load the tidyverse.

```{r}
library(tidyverse)
```

And read in the attendance data:

```{r}
attendance <- read_csv("https://dwillis.github.io/sports-data-files/attendance.csv")
```

Using the tidyverse's `pivot_longer` function, turn the currently wide data into longer data and save it as a new dataframe called `attendance_longer`:


```{r}
attendance_longer <- attendance |> pivot_longer(cols = starts_with("20"), names_to = "Year", values_to = "Attendance")
```

Let's try to figure out whether football attendance has returned to "normal" by 2024, defining "normal" as 2019 figures. For that we'll need to isolate those two years using `filter()` and get rid of any NAs:

```{r}
attendance_two_years <- attendance_longer |> filter(Year == 2019 | Year == 2024) |> na.omit()
```

Now, in order to calculate a percentage change between those two years, we'll need to go back to wider data and again ditch teams where we don't have both years:

```{r}
attendance_wide <- attendance_two_years |> pivot_wider(names_from = Year, values_from = Attendance) |> na.omit()
```

Finally, we'll add a percentage change column to that dataframe (remember, it's (New-Old)/Old) and arrange it to see the results.

```{r}
attendance_wide <- attendance_wide |> mutate(percent_change = (`2024` -`2019`)/`2019` * 100) |> arrange(percent_change)
```

Some of these clearly are influenced by performance on the field. Below, describe the most interesting results (positive or negative) and what you think might be going on with them. Feel free to do some googling and describe what other data would be helpful to clarify what the best stories here might be. What surprises you about these results?

**ANSWER**
As poorly has Maryland has played, their numbers actually increased by 11% with a (3-9 record in 2019 and 4-8 in 2024). That is suprising to me and makes me wonder why fans are still showing up? Possibly joining the Big Ten conference brings in new opponents that may draw a large  croud regardless of Maryland's record. The schools close location to D.C. might make it easier to access for alumni and fans. Game day culture at UMD has become big with crowded student sections and tailgaiting traditions. A good story angle would be that even struggling programs attract fans who believe the team might improve under new players 

